Proposal outline/notes

### 2012-06-13 docs/Jesse/Proposal/outline.tex
adding dg to proposal

An additional reason for the popularity of DG methods is that they can be interpreted as stabilized FE methods through appropriate choices of the numerical flux \todo{add Brezzi citation}. We will illustrate this with the pure convection equation in 1D.

Local conservation: in 1D, numerical solution at outflow is exact.

@@

For example, for the pure convection problem, the solution has only regularity in the streamline direction, but is only $L^2$ in the crosswind direction. As a consequence, the boundary trace of the solution is defined only in the direction of convection. The upwind DG method addresses the above issue by choosing the numerical flux to be the upwind flux; in this case, the DG numerical flux can be viewed as imparting additional regularity than is present in the solution \cite{DPG1,DPG3}.

@@

\bibliographystyle{plain}
\bibliography{CFD_intro,DPG_old,paper,NS}

>>>>

An additional reason for the popularity of DG methods is that they can be interpreted as stabilized FE methods through appropriate choices of the numerical flux \cite{Brezzi20063293}. We will illustrate this with the pure convection equation in 1D ---
\begin{align*}
\pd{\left(\beta u\right)}{x} &= f\\
u(0) &= u_0.
\end{align*}
The DG formulation is derived by multiplying by a test function $v$ with support only on a single element $K = [x_K,x_{K+1}]$ and integrating by parts. The boundary term is left alone, such that the local formulation is 
\[
\left.\beta u v\right|_{x_K}^{x_{K+1}} + \int_K -\beta u \pd{v}{x} = \int_K f v,
\]
and the global formulation is recovered by summing up all element-wise local formulations. However, the boundary term in the local formulation is presently ill-defined, as both $u$ and $v$ are dual-valued over element boundaries. Consequently, we make the choice to define the values of $u$ on the boundary (the \emph{traces} of $u$) as
\begin{align*}
u(x_K) \coloneqq u(x_K^-), \quad u(x_{K+1}) \coloneqq u(x_{K+1}^-),
\end{align*}
where $u(x_K^-)$ is the value of $u$ at $x_K$ as seen from the left, and $u(x_K^+)$ the value as seen from the right. Similarly, the traces of $v$ are defined to be
\begin{align*}
v(x_K) \coloneqq v(x_K^+), \quad v(x_{K+1}) \coloneqq v(x_{K+1}^-),
\end{align*}
Notice the specific choice of $v(x_K)$ --- we have chosen the value of $v(x_K)$ to be the value in the upwind direction.

@@

For example, for the pure convection problem, the solution has a distributional derivative in the streamline direction, but is only $L^2$ in the crosswind direction. As a consequence of the regularity of the solution, the boundary trace of the solution is defined only in the direction of convection. The upwind DG method addresses the above issue by choosing the numerical flux to be the upwind flux; in this case, the DG numerical flux can be viewed as imparting additional regularity than is present in the solution \cite{DPG1,DPG3}.

@@

\bibliographystyle{plain}
\bibliography{CFD_intro,DPG_old,paper,LitRev}
